Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-08 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
Section III

ParSpec is such an observable-based
\sout{In the ParSpec expansion this is accomplished by performing a} bivariate
expansion of Eq.~\eqref{eq:general_deviation} in the detector-frame BH's mass
$M$ and spin $\chi$, \sout{namely}
%
\begin{subequations}
\label{eq:kerr_expansion}
\begin{align}
\omega_{k} &= \frac{1}{M} \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \omega^{(n)}_{k} \, \left( 1 + \gamma \delta \omega^{(n)}_{k} \right), \\
\tau_{k}   &= M     \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \tau^{(n)}_{k}   \, \left( 1 + \gamma \delta \tau^{(n)}_{k} \right),
\end{align}
\end{subequations}
%
where $\omega_{k}^{(n)}$ and $\tau_{k}^{(n)}$ are dimensionless coefficients of the
expansion in spin for Kerr QNMs;
%
$\delta\omega_{k}^{(n)}$ and $\delta\tau_{k}^{(n)}$ are source-independent
dimensionless coefficients that characterize the corrections to the Kerr QNM at
each spin-order;
%
and $N_{\rm max}$ is the order of the spin expansion.
%
All source dependence on a given modified gravity theory is contained in the
dimensionless parameter $\gamma$, given by
%
\begin{equation}
\gamma
%
= \left(\frac{\ell}{M_{\rm s}}\right)^{p}
%
= \left[
\frac{\ell c^2 (1 + z)}{G M}
\right]^{p},
\label{eq:def_gamma}
\end{equation}
%
which depends on a parameter $\ell$, with dimensions of length (and sets the
length-scale at which beyond-GR modifications become important), and
the exponent $p$ which is related to how the beyond-GR modifications are
added to the Einstein-Hilbert action.
%
In Eq.~\eqref{eq:def_gamma} we made \sout{$\ell$} \ag{$\gamma$} dimensionless  by using the other
length-scale associated with the remnant black hole, i.e., its source-frame mass
$M_{\rm s}$, which we can also write in terms of the detector-frame mass $M$ through
the redshift $z$~\cite{Krolak:1987ofj}. \ag{Also, dividing by the factor $G/c^2$ allows us to
express $\ell$ in physically intuitive metric units.}

@@

The GW signal from a BBH merger can be determined in GR by a unique set of
parameters $\bm{\theta}$, that includes the masses and spins of the two BHs,
$(m_1,\, m_2,\, \vec{s}_1,\, \vec{s}_2)$, the sky location determined by the
luminosity distance $d_L$, right ascension $\alpha$ and declination $\delta$,
and the orientation of the binary given by the inclination $\iota$ and polarization $\psi$
angles.
%
The set is completed by the choice of a reference time $t_0$ and phase
$\phi_0$. If we further assume that the spins of the individual BHs are
restricted to be parallel to the orbital angular momentum, we reduce the six
components of spin to just two, $(\chi_1, \chi_2)$ and our entire parameter set
from 15 to 11.
%
Let us also define some additional parameters and set some conventions that
will be useful in our analysis later, namely, the total mass
%
$M=m_1+m_2$,
%
the chirp mass
%
$\mathcal {M}=(m_{1}m_{2})^{3/5}/(m_{1}+m_{2})^{1/5}$,
%
the asymmetric mass ratio $q=m_1/m_2$, with the convention $m_1 \geqslant m_2$ (thus $q \geqslant 1$),
and the symmetric mass ratio of the binary, $\nu = m_1m_2/(m_1+m_2)^2$.

The polarization of the GW signal (in the observer's frame) is,
%
\begin{equation}
h_+(\iota,\varphi_0;t ) - i h_\times(\iota,\varphi_0;t) = \sum_{l, m} {}_{-\!2}Y_{l m}(\iota,\varphi_0)\, h_{l m}(t)\,,
\end{equation}
%
where ${}_{-\!2}Y_{l m}(\iota,\varphi_0)$ are the $-2$ spin-weighted spherical
harmonics described by standard angular dependence indices $l$ and $m$.
%
% --------------------------------
% HS: maybe not necessary anymore:
% \footnote{Note that we use $l$ to denote the angular dependence index and $\ell$ to denote the coupling constant in Sec.~\ref{sec:review_parspec}.}
% --------------------------------
%
As our baseline model, that is, the GR model upon which non-GR modifications will be added to, we use the computationally efficient (time-domain) multipolar waveform model for quasicircular spin-aligned BBH mergers described in Ref.~\cite{Mihaylov:2021bpf} (henceforth referred to as \SEOB{}\footnote{This waveform model is available in \texttt{LALSuite}~\cite{lalsuite} as the \texttt{SEOBNRv4HM\_PA} waveform approximant.}) which contains the modes, $(l, |m|)=(2,2),(2,1)$, $(3,3)$, $(4,4)$, and $(5,5)$~\cite{Cotesta:2018fcv,Mihaylov:2021bpf}.
%
The model uses an effective-one-body (EOB) approach combined with a
post-adiabatic solution of the equations of motion to describe the
inspiral-plunge waveform, $h_{l m}^\mathrm{insp-plunge}$.
%
An accurate description of the merger is incorporated through calibration with
NR simulations, as described in Ref.~\cite{Cotesta:2018fcv}, along with
information for the merger and ringdown phases, from BH perturbation theory.
The merger-ringdown waveform, $h_{l m}^\textrm{merger-RD}$, is then stitched to
inspiral-plunge waveform, $h_{l m}^\textrm{insp-plunge}$ at a certain time $t = t^{\textrm{match}}_{l m}$, as
%
\begin{equation}
h_{l m}(t) =
h_{l m}^\textrm{insp-plunge}\, \theta(t_\textrm{match}^{l m} - t)
+ h_{l m}^\textrm{merger-RD}\,\theta(t-t_\textrm{match}^{l m})\,,
\end{equation}
%
where $\theta(t)$ is the Heaviside step function.
%
The merger-ringdown waveform is expressed as an exponentially damped
sinusoid~\cite{Bohe:2016gbl,Cotesta:2018fcv,Mihaylov:2021bpf}
%
\begin{equation}
\label{RD}
h_{l m}^{\textrm{merger-RD}}(t) = \nu \ \tilde{A}_{l m}(t)\ e^{i \tilde{\phi}_{l m}(t)} \ e^{-i \sigma_{l m 0}(t-t^{\textrm{match}}_{l m})}\,,
\end{equation}
%
where
%
\begin{align}
\sigma_{l m 0} = {\rm Re}(\sigma_{l m 0}) + i \, {\rm Im}(\sigma_{l m 0})
= \omega_{l m 0} - i / \tau_{l m 0}\,,
\end{align}
%
are the complex frequencies of the fundamental ($n=0$ overtone) QNMs of the remnant BH.
%
The frequencies $\omega_{l m 0}$ and damping times $\tau_{l m 0}$ of the $(l,
m)$ QNM can be read off from the real and imaginary parts of $\sigma_{l m 0}$
respectively.
%
The functions $\tilde{A}_{l m}(t)$ and $\tilde{\phi}_{l m}(t)$ are defined
in Ref.~\cite{Bohe:2016gbl,Cotesta:2018fcv}.

In the $\SEOB$ model~\cite{Mihaylov:2021bpf}, the complex frequencies
$\sigma_{l m 0}$ are computed by first determining the final mass and spin from
estimates of the initial masses and spins through NR-fitting-formulas~\cite{Taracchini:2013rva,Hofmann:2016yih}
and then converting them to the complex frequencies using BH perturbation
theory-inspiral analytical fits~\cite{Berti:2005ys,Berti:2009kk}.
%
Hence,
%
\begin{subequations}
\begin{align}
\omega_{l m 0}^{\text{GR}} &= \omega_{l m 0}^{\text{GR}}(m_1, m_2, \chi_1, \chi_2)\,,
\\
\tau _{l m 0}^{\text{GR}} &= \tau _{l m 0}^{\text{GR}}(m_1, m_2, \chi_1, \chi_2)\,,
\end{align}
\end{subequations}
%
where $(\omega_{l m 0}^{\text{GR}}, \tau_{l m 0}^{\text{GR}} )$ refer to the
QNM predictions in baseline \SEOB{} model.

In this paper, however, we introduce a completely different expression of the
QNM frequencies $(\omega_{l m 0}, \tau_{l m 0})$, as a series expansion on the
spin of the remnant object $\chi_f$ as we described in Sec.~\ref{sec:review_parspec}; cf.~Eqs.~\eqref{eq:kerr_expansion}.
%
For the rest of the paper, in keeping with the conventions introduced in
Sec.~\ref{sec:review_parspec}, we are going to use the index $k$ to refer to
$(l, m,n=0)$ mode. Accordingly, we are going to use $(\omega_k, \tau _k)$ to
refer to $(\omega_{l m 0}, \tau _{l m 0})$ respectively.

Following the ParSpec expansion, our QNM frequencies are expressed through functions,
%
\begin{subequations}
\begin{align}
\omega_k &= \omega_k(m_1, m_2, \chi_1, \chi_2,\ell, \{\delta \omega_k^{(j)}\}),\\
\tau_k   &= \tau _k(m_1, m_2, \chi_1, \chi_2, \ell, \{\delta \tau_k^{(j)}\}),
\end{align}
\end{subequations}
%
where we fixed a certain value of $p$, and the estimates of $(m_1,\, m_2,\, \chi_1,\, \chi_1)$ are used to predict
the final BH mass $M_f$ and spin $\chi_f$ through~\cite{Taracchini:2013rva,Hofmann:2016yih}.
%
Additionally, both $\omega_k$ and $\tau_k$ depend on the coupling constant $\ell$
defined in Eq.~\eqref{eq:def_gamma} as well as the sets, $\{\delta
\omega_k^{(n)}\},$ and $\{\delta \tau_k^{(n)}\}$, where $n$ ranges from zero to
the power of spin in Eqs.~\eqref{eq:kerr_expansion} up to which we include
corrections.
%
For example, if we restrict ourselves to corrections with $n=0$, or
$\chi_f^0$, to the $k=(2,2,0)$ mode only, then we would just have two
additional parameters, $\{\delta \omega_0^{(0)}, \delta \tau_0^{(0)}\}$.
%
In this paper, we will restrict ourselves to corrections to the $n=0$ and
$n=1$ terms, that is, only up to the leading-order in spin $\chi_f$ term.

@@

Let us now establish the connection between theory-agnostic framework of the
$\pSEOB$ waveform model and the theory-specific QNM calculations.
%
We begin by rewriting $\omega_{k}$ and $\tau_{k}$, given by Eqs.~\eqref{eq:kerr_expansion} as,
%
\begin{subequations}
\label{eq:qnm_nongr_iso}
\begin{align}
    M \omega_{k} &= \gamma \left( \delta\omega^{(0)}_{k} \omega^{(0)}_{k} + \chi \delta\omega^{(1)}_{k} \omega^{(1)}_{k} \right)
    + \sum_{n=0}^{N_{\rm max}} \chi^{n} \omega^{(n)}_{k} \,,
\\
    \tau_{k}/M   &= \gamma \left( \delta\tau^{(0)}_{k} \tau^{(0)}_{k} + \chi \delta\tau^{(1)}_{k} \tau^{(1)}_{k} \right)
    + \sum_{n=0}^{N_{\rm max}} \chi^{n} \tau^{(n)}_{k}\,,
\end{align}
\end{subequations}
%
We will work exclusively with the $\ell = m = 2$ and $n = 0$ (fundamental mode)
QNM, i.e., $k = \{2,\, 2,\, 0\}$. For this reason we will omit the subcripts
$k$ hereafter and no ambiguity in notation between the overtone index and the
subscript $n$ in the series should arise.
%
We have also pulled out from the sum all beyond-GR corrections, restricting
ourselves to the nonspinning ($n=0$) and linear-order in spin ($n=1$)
corrections to the Kerr QNMs.
%
% These corrections are encoded in the coefficients $\delta\omega^{(n)}$ and $\delta\tau^{(n)}$.

@@

In the end, the $\pSEOB$ waveform model with theory-specific informed
QNMs has only $\ell$ as a free beyond-GR parameter.
%
We emphasize that our procedure is different from that of
Ref.~\cite{Carullo:2021dui} which, for a given value of $p$, varied all $\ell$,
$\delta\omega^{(n)}$, and $\delta\tau^{(n)}$ parameters, and then proceeded
to use the posteriors on $\ell$ to draw conclusions to any theory that predict
corrections at a given $p$.
%
This is a critical difference between our work and Ref.~\cite{Carullo:2021dui}.
In particular, as we will see in Sec.~\ref{sec:results}, adding theory-specific
information to the ParSpec expansion coefficients \emph{can lead to different interpretations
of the final bounds on $\ell$, even for different theories which predict the same value of the exponent $p$.}

\begin{table}[t]
\begin{tabular}{c | c c c c c c}
\hline
\hline
Theory & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & Ref.  \\
\hline
EdGB        & 4 & 0.0107    & 0.0044    & $-$0.2480 & $-$1.101 &  \cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} \\
dCS         & 4 & 3.1964    & 6.3619    & 41.199    & 794.66   &  \cite{Wagle:2021tam}  \\
cubic EFT   & 4 & $-$0.5813 & 2.6469    & $-$3.8620 & 265.12   & \cite{Cano:2021myl} \\
quartic EFT & 6 & $-$0.2114 & $-$0.6070 & $-$1.5263 & 171.35   & \cite{Cano:2021myl}  \\
\hline
\hline
\end{tabular}
\caption{Summary of theory-specific QNM calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(n)}_{220}$ and decay time $\delta \tau^{(n)}_{220}$, and the
references from which we used the results from.
}
\label{tab:ref_theories_qnms}
\end{table}

As we have seen in Sec.~\ref{sec:review_theories}, the QNM of slowly-rotating
BHs is modified gravity theories come in two families depending on their
transformation upon a parity transformation: axial and polar. Which is the one
we use to match against Eqs.~\eqref{eq:qnm_nongr_iso}?
%
To answer this question one has to work with a chosen theory and perform a
translation between the metric perturbations $h_{\mu\nu}$ in the
Regge-Wheeler-Zerilli gauge~\cite{Regge:1957td,Zerilli:1970se} and connect it with the transverse-traceless gauge
used to described GWs (see e.g.~Ref.~\cite{Maggiore:2018sht}, Chapter~12),
%
In GR, both axial and polar QNMs are isospectral and hence which QNM we use to
model the ringdown makes no difference.
%
In beyond-GR theories, isospectrality is in general broken (see
in Ref.~\cite{Hui:2021cpm} for a counterexample).
%
Thus, how axial and polar gravitational QNMs appear in the GW signal has to be
answered on a theory-by-theory basis.
%
This is outside the scope of this paper and here we take the more pragmatic
approach of simply choosing the \emph{least damped gravitational mode between
the two parities.}
%
The justification is this is the mode (if excited during the merger) which is
the most likely to appear in the signal.

We performed the mapping between theory-specific QNM calculation and the ParSpec framework
under the hypothesis above for the theories listed in Sec.~\ref{sec:review_theories}.
%
We summarize our result in Table~\ref{tab:ref_theories_qnms} and
leave the details of our calculations to Appendix~\ref{app:map_details}.

In Fig.~\ref{fig:example_waveform} we show an illustrative waveform for GR (solid line; using the \texttt{SEOBNR} model) and in the cubic EFT of gravity (dashed line; using the \pSEOB{} model),
including the leading-order $n = 0$ deformations to the fundamental QNM (see Table~\ref{tab:ref_theories_qnms}).
%
We chose binary parameters similar to GW150914, consisting of a quasicircular,
nonspinning binary with detector-frame masses
$m_1^{\rm dec} = 39\msun$ and $m_2^{\rm dec} = 31\msun$.
%
The top panel shows the $+$ polarization of the GW $h_{+}$ in both theories,
while the bottom panel shows the amplitude $|h| = (h_{+}^2 + h_{\times}^2)^{1/2}$
and instantaneous frequency $f$, all as functions of time $t$.
%
The signal are identical up to the merger, after which they differ during the ringdown.
%
By construction, the ringdown is lasts longer for the cubic EFT of GR waveform and
with a smaller instantaneous frequency (see the bottom panel) due to the negative value of the
$\delta\omega^{(0)}$ coefficient in this theory.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/example_waveform_cubicEFT.pdf}
\caption{(Color online).~Example GW signal with GW1509140-like parameters
for both GR (solid line) and cubic EFT of GR (dashed line) with the leading-order
$n=0$ modifications to the fundamental QNM, for $\ell_{\rm cEFT} = 65$~km.
%
The former is computed with the \SEOB{} model, while the latter with the
\pSEOB{} model, with ringdown modifications according to the results in
Table~\ref{tab:ref_theories_qnms}.
%
Top panel: the $+$ polarization $h_{+}(t)$. Bottom panel: the GW amplitude
$|\tilde{h}(t)|$ (left axes) and the instantaneous frequency $f(t)$ (right axes).
}
\label{fig:example_waveform}
\end{figure}

>>>>

ParSpec is such an observable-based
\sout{In the ParSpec expansion this is accomplished by performing a} bivariate
expansion of Eq.~\eqref{eq:general_deviation} in the detector-frame BH's mass
$M$ and spin $\chi$, \sout{namely}
%
\begin{subequations}
\label{eq:kerr_expansion}
\begin{align}
\omega_{k} &= \frac{1}{M} \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \omega^{(n)}_{k} \, \left( 1 + \gamma \delta \omega^{(n)}_{k} \right), \\
\tau_{k}   &= M     \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \tau^{(n)}_{k}   \, \left( 1 + \gamma \delta \tau^{(n)}_{k} \right),
\end{align}
\end{subequations}
%
where $\omega_{k}^{(n)}$ and $\tau_{k}^{(n)}$ are dimensionless coefficients of the
expansion in spin for Kerr QNMs\footnote{\ag{From this point onwards, to prevent ambiguity,
we will use $n$ to denote the series expansion index, and not the spherical 
harmonic overtone.}};
%
$\delta\omega_{k}^{(n)}$ and $\delta\tau_{k}^{(n)}$ are source-independent
dimensionless coefficients that characterize the corrections to the Kerr QNM at
each spin-order;
%
and $N_{\rm max}$ is the order of the spin expansion.
%
All source dependence on a given modified gravity theory is contained in the
dimensionless parameter $\gamma$, given by
%
\begin{equation}
\gamma
%
= \left(\frac{\ell}{M_{\rm s}}\right)^{p}
%
= \left[
\frac{\ell c^2 (1 + z)}{G M}
\right]^{p},
\label{eq:def_gamma}
\end{equation}
%
which depends on a parameter $\ell$, with dimensions of length (and sets the
length-scale at which beyond-GR modifications become important), and
the exponent $p$ which is related to how the beyond-GR modifications are
added to the Einstein-Hilbert action.
%
In Eq.~\eqref{eq:def_gamma} we made \sout{$\ell$} \ag{$\gamma$} dimensionless  by using the other
length-scale associated with the remnant black hole, i.e., its source-frame mass
$M_{\rm s}$\footnote{\ag{Throughout this paper, we maintain the convention of 
using the subscript ``s" to denote source-frame quantities and plain baseline 
symbols for detector-frame measurements.}} which we can also write in terms of the detector-frame mass $M$ through
the redshift $z$~\cite{Krolak:1987ofj}. \ag{Also, dividing by the factor $G/c^2$ allows us to
express $\ell$ in physically intuitive metric units.}

@@

The GW signal from a BBH merger can be determined in GR by a unique set of
parameters $\bm{\theta}$, that includes the masses and spins of the two BHs,
$(m_1,\, m_2,\, \vec{s}_1,\, \vec{s}_2)$, the sky location determined by the
luminosity distance $d_L$, right ascension $\alpha$ and declination $\delta$,
and the orientation of the binary given by the inclination $\iota$ and polarization $\psi$
angles.
%
The set is completed by the choice of a reference time $t_0$ and phase
$\phi_0$. If we further assume that the spins of the individual BHs are
restricted to be parallel to the orbital angular momentum, we reduce the six
components of spin to just two, $(\chi_1, \chi_2)$ and our entire parameter set
from 15 to 11.
%
Let us also define some additional parameters and set some conventions that
will be useful in our analysis later, namely, the total mass
%
$M=m_1+m_2$,
%
the chirp mass
%
$\mathcal {M}=(m_{1}m_{2})^{3/5}/(m_{1}+m_{2})^{1/5}$,
%
the asymmetric mass ratio $q=m_1/m_2$, with the convention $m_1 \geqslant m_2$ (and thus $q \geqslant 1$),
and the symmetric mass ratio of the binary, $\nu = m_1m_2/(m_1+m_2)^2$.

The polarization of the GW signal (in the observer's frame) is,
%
\begin{equation}
h_+(\iota,\varphi_0;t ) - i h_\times(\iota,\varphi_0;t) = \sum_{l, m} {}_{-\!2}Y_{l m}(\iota,\varphi_0)\, h_{l m}(t)\,,
\end{equation}
%
where ${}_{-\!2}Y_{l m}(\iota,\varphi_0)$ are the $-2$ spin-weighted spherical
harmonics.
%
% --------------------------------
% HS: maybe not necessary anymore:
% \footnote{Note that we use $l$ to denote the angular dependence index and $\ell$ to denote the coupling constant in Sec.~\ref{sec:review_parspec}.}
% --------------------------------
%
As our baseline model, that is, the GR model upon which non-GR modifications 
will be added, we use the computationally efficient (time-domain) multipolar 
waveform model for quasicircular spin-aligned BBH mergers described in 
Ref.~\cite{Mihaylov:2021bpf} (henceforth referred to as \SEOB{}\footnote{This waveform 
model is available in \texttt{LALSuite}~\cite{lalsuite} as the \texttt{SEOBNRv4HM\_PA} 
waveform approximant.}) which contains the modes, $(l, |m|)=(2,2),(2,1)$, $(3,3)$, $(4,4)$, 
and $(5,5)$~\cite{Cotesta:2018fcv,Mihaylov:2021bpf}.
%
The model uses an effective-one-body (EOB) approach combined with a
post-adiabatic solution of the equations of motion to describe the
inspiral-plunge waveform, $h_{l m}^\mathrm{insp-plunge}$.
%
An accurate description of the merger is incorporated through calibration with
NR simulations, as described in Ref.~\cite{Cotesta:2018fcv}, along with
information for the merger and ringdown phases, from BH perturbation theory.
The merger-ringdown waveform, $h_{l m}^\textrm{merger-RD}$, is then stitched to
inspiral-plunge waveform, $h_{l m}^\textrm{insp-plunge}$ at a certain time $t = t^{\textrm{match}}_{l m}$, as
%
\begin{equation}
h_{l m}(t) =
h_{l m}^\textrm{insp-plunge}\, \Theta(t_\textrm{match}^{l m} - t)
+ h_{l m}^\textrm{merger-RD}\,\Theta(t-t_\textrm{match}^{l m})\,,
\end{equation}
%
where $\Theta(t)$ is the Heaviside step function \agcomm{Changed $\theta(t)$ to $\Theta(t)$,
because $\theta$ was already used to denote the parameter space.}.
%
The merger-ringdown waveform is expressed as an exponentially damped
sinusoid~\cite{Bohe:2016gbl,Cotesta:2018fcv,Mihaylov:2021bpf}
%
\begin{equation}
\label{RD}
h_{l m}^{\textrm{merger-RD}}(t) = \nu \ \tilde{A}_{l m}(t)\ e^{i \tilde{\phi}_{l m}(t)} \ e^{-i \sigma_{l m 0}(t-t^{\textrm{match}}_{l m})}\,,
\end{equation}
%
where
%
\begin{align}
\sigma_{l m 0} = {\rm Re}(\sigma_{l m 0}) + i \, {\rm Im}(\sigma_{l m 0})
= \omega_{l m 0} - i / \tau_{l m 0}\,,
\end{align}
%
are the complex frequencies of the fundamental ($0$-th overtone) QNMs of the remnant BH.
%
\ag{Going back to the convention established above to use the index $k$ to now denote the $(l, m,0)$
mode,  the frequencies $\omega_k$ and damping times $\tau_k$ can be read off from the 
real and imaginary parts of $\sigma_k$ respectively.}
%
The functions $\tilde{A}_{l m}(t)$ and $\tilde{\phi}_{l m}(t)$ are defined
in Ref.~\cite{Bohe:2016gbl,Cotesta:2018fcv}.

In the $\SEOB$ model~\cite{Mihaylov:2021bpf}, the complex frequencies
$\sigma_k$ are computed by first determining the final mass and spin from
estimates of the initial masses and spins through NR-fitting-formulas~\cite{Taracchini:2013rva,Hofmann:2016yih}
and then converting them to the complex frequencies using BH perturbation
theory-inspired analytical fits~\cite{Berti:2005ys,Berti:2009kk}.
%
Hence,
%
\begin{subequations}
\begin{align}
\omega_k^{\text{GR}} &= \omega_k^{\text{GR}}(m_1, m_2, \chi_1, \chi_2)\,,
\\
\tau _k^{\text{GR}} &= \tau _k^{\text{GR}}(m_1, m_2, \chi_1, \chi_2)\,,
\end{align}
\end{subequations}
%
where $(\omega_k^{\text{GR}}, \tau_k^{\text{GR}} )$ refer to the
GR QNM predictions in the baseline \SEOB{} model.

\ag{In this paper, we replace these GR predictions with QNM 
frequencies defined through the ParSpec expansion introduced in 
Sec.~\ref{sec:review_parspec}; cf.~Eqs.~\eqref{eq:kerr_expansion}. Hence,}
%
\begin{subequations}
\begin{align}
\omega_k &= \omega_k(m_1, m_2, \chi_1, \chi_2,\ell, \{\delta \omega_k^{(n)}\}),\\
\tau_k   &= \tau _k(m_1, m_2, \chi_1, \chi_2, \ell, \{\delta \tau_k^{(n)}\}),
\end{align}
\end{subequations}
%
\ag{where the mass and spin of the remnant object $(M_f,\chi_f)$ are themselves functions of
$(m_1,\, m_2,\, \chi_1,\, \chi_2)$ (through~\cite{Taracchini:2013rva,Hofmann:2016yih}), and we fix $p$
to a certain theory-specific value. Additionally,  the frequencies depend on the ParSpec coefficients $\{\delta
\omega_k^{(n)},\delta \tau_k^{(n)}\}$,  and the coupling constant $\ell$.}

\sout{For example, if we restrict ourselves to corrections with $n=0$, or
$\chi_f^0$, to the $k=(2,2,0)$ mode only, then we would just have two
additional parameters, $\{\delta \omega_0^{(0)}, \delta \tau_0^{(0)}\}$.
%
In this paper, we will restrict ourselves to corrections to the $n=0$ and
$n=1$ terms, that is, only up to the leading-order in spin $\chi_f$ term.}

@@

Let us now establish the connection between the theory-agnostic \agcomm{or independent?} framework of the
$\pSEOB$ waveform model and the theory-specific QNM calculations.  \ag{In this paper we will restrict ourselves to
corrections to the leading and next-to-leading order terms in the ParSpec expansion, as well as the fundamental 
$k = \{2,\, 2,\, 0\}$ QNM. For this reason,  we will omit the subscripts $k$ hereafter 
and rewrite $\omega$ and $\tau$, given by} Eqs.~\eqref{eq:kerr_expansion} as,
%
\begin{subequations}
\label{eq:qnm_nongr_iso}
\begin{align}
    M \omega &= \gamma \left( \delta\omega^{(0)} \omega^{(0)} + \chi \delta\omega^{(1)} \omega^{(1)} \right)
    + \sum_{n=0}^{N_{\rm max}} \chi^{n} \omega^{(n)} \,,
\\
    \tau/M   &= \gamma \left( \delta\tau^{(0)} \tau^{(0)} + \chi \delta\tau^{(1)} \tau^{(1)} \right)
    + \sum_{n=0}^{N_{\rm max}} \chi^{n} \tau^{(n)}\,,
\end{align}
\end{subequations}
%
\sout{We will work exclusively with the $\ell = m = 2$ and $n = 0$ (fundamental mode)
QNM, i.e., $k = \{2,\, 2,\, 0\}$. For this reason we will omit the subcripts
$k$ hereafter and no ambiguity in notation between the overtone index and the
subscript $n$ in the series should arise.}
%
We have \sout{also} pulled out from the sum all beyond-GR corrections, restricting
ourselves to the nonspinning ($n=0$) and linear-order in spin ($n=1$)
corrections to the Kerr QNMs.
%
% These corrections are encoded in the coefficients $\delta\omega^{(n)}$ and $\delta\tau^{(n)}$.

@@

In the end, the $\pSEOB$ waveform model with theory-specific
QNMs has only $\ell$ as a free beyond-GR parameter.
%
We emphasize that our procedure is different from that of
Ref.~\cite{Carullo:2021dui} which, for a given value of $p$, varied all $\ell$,
$\delta\omega^{(n)}$, and $\delta\tau^{(n)}$ parameters, and then proceeded
to use the posteriors on $\ell$ to draw conclusions to any theory that predict
corrections at a given $p$.
%
This is a critical difference between our work and Ref.~\cite{Carullo:2021dui}.
In particular, as we will see in Sec.~\ref{sec:results}, adding theory-specific
information to the ParSpec expansion coefficients \emph{can lead to different interpretations
of the final bounds on $\ell$, even for different theories which predict the same value of the exponent $p$.}

\begin{table}[t]
\begin{tabular}{c | c c c c c c}
\hline
\hline
Theory & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & Ref.  \\
\hline
EdGB        & 4 & 0.0107    & 0.0044    & $-$0.2480 & $-$1.101 &  \cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} \\
dCS         & 4 & 3.1964    & 6.3619    & 41.199    & 794.66   &  \cite{Wagle:2021tam}  \\
cubic EFT   & 4 & $-$0.5813 & 2.6469    & $-$3.8620 & 265.12   & \cite{Cano:2021myl} \\
quartic EFT & 6 & $-$0.2114 & $-$0.6070 & $-$1.5263 & 171.35   & \cite{Cano:2021myl}  \\
\hline
\hline
\end{tabular}
\caption{Summary of theory-specific QNM calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(n)}_{220}$ and decay time $\delta \tau^{(n)}_{220}$, and the
references from which we used the results from.\agcomm{Just leaving this comment 
as a reminder to expand this table to include the GR values from Maselli et al.}
}
\label{tab:ref_theories_qnms}
\end{table}

As we have seen in Sec.~\ref{sec:review_theories}, QNMs of slowly-rotating
BHs in modified gravity theories can belong to two families depending on how 
they behave under a parity transformation: axial and polar. Which one do
we use to match with Eqs.~\eqref{eq:qnm_nongr_iso}?
%
To answer this question one has to work with a chosen theory and perform a
translation between the metric perturbations $h_{\mu\nu}$ in the
Regge-Wheeler-Zerilli gauge~\cite{Regge:1957td,Zerilli:1970se} and connect it with the transverse-traceless gauge
used to described GWs (see e.g.~Ref.~\cite{Maggiore:2018sht}, Chapter~12),
%
In GR, both axial and polar QNMs are isospectral and hence which QNM we use to
model the ringdown makes no difference.
%
In beyond-GR theories, isospectrality is in general broken (see
in Ref.~\cite{Hui:2021cpm} for a counterexample).
%
Thus, how axial and polar gravitational QNMs appear in the GW signal has to be
answered on a theory-by-theory basis.
%
This is outside the scope of this paper and here we take the more pragmatic
approach of simply choosing the \emph{least damped gravitational mode between
the two parities.}
%
The justification is this is the mode (if excited during the merger) which is
the most likely to appear in the signal.
We performed the mapping between theory-specific QNM calculations and the ParSpec framework
under the hypothesis above, for the theories listed in Sec.~\ref{sec:review_theories}.
%
We summarize our results in Table~\ref{tab:ref_theories_qnms} and
leave the details of our calculations to Appendix~\ref{app:map_details}.

In Fig.~\ref{fig:example_waveform} we show an illustrative waveform for GR (solid line; using the \texttt{SEOBNR} model) and in the cubic EFT of gravity (dashed line; using the \pSEOB{} model),
including the leading-order $n = 0$ deformations to the fundamental QNM (see Table~\ref{tab:ref_theories_qnms}).
%
We chose binary parameters similar to GW150914,  but non-spinning, with detector-frame masses
\sout{$m_1^{\rm dec} = 39\msun$ and $m_2^{\rm dec} = 31\msun$}
\ag{$m_1 = 39\msun$ and $m_2 = 31\msun$} \agcomm{My suggestion would be to follow what
I have written in footnote 2, and use unsubscripted symbols to denote detector frame quantities throughout.}.
%
The top panel shows the $+$ GW polarization $h_{+}$ in both theories,
while the bottom panel shows the amplitude $|h| = (h_{+}^2 + h_{\times}^2)^{1/2}$
and instantaneous frequency $f$, all as functions of time $t$.
%
The signals are identical up to the merger,  \ag{specifically, the time $t_\textrm{match}$ 
defined in Sec.~\ref{sec:review_pSEOB}}, after which they differ during the ringdown.
%
By construction, the ringdown lasts longer for the cubic EFT of GR waveform and
with a smaller instantaneous frequency (see the bottom panel) due to the negative value of the
$\delta\omega^{(0)}$ coefficient in this theory.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/example_waveform_cubicEFT.pdf}
\caption{(Color online).~Example GW signal with GW150914-like parameters
for both GR (solid line) and cubic EFT of GR (dashed line) with the leading-order
$n=0$ modifications to the fundamental QNM, for $\ell_{\rm cEFT} = 65$~km.
%
The former is computed with the \SEOB{} model, while the latter with the
\pSEOB{} model, with ringdown modifications according to the results in
Table~\ref{tab:ref_theories_qnms}.
%
Top panel: the $+$ polarization $h_{+}(t)$. Bottom panel: the GW amplitude
$|\tilde{h}(t)|$ (left axes) and the instantaneous frequency $f(t)$ (right axes).
}
\label{fig:example_waveform}
\end{figure}
